CDSR Human Data/Modelling Paper

### 2013-09-09 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK updated the text introducing the likert scale results and the comparision of the sweetspot across the flipped trials

All three trial types use the same visual stimuli and accompanying spatial descriptions. Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment. The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and the military engagement stimuli were always accompanied by the description \emph{safety}.

@@

\subsection{Participants}

43 undergraduate students from two different third level institutes took part in the study.

@@

The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region. For each stimuli domain (classroom, map, military engagement) three points were pre-selected such that for each particular stimuli one of the points was located near the expected prototypical center of the described spatial region and the other two points were at varying distances from this location. For clarity it is worth emphaising that, although the same three points were used for all the stimuli from a particular domain, the particular point that acted as the sweetspot for a particular stimuli within a domain set would vary from stimuli to stimuli. The selection of these points was informed by the results of a pilot study. Figure \ref{fig:locations} illustrates the points selected for each domain and overlays them on one of the visual stimuli from each domain. Table \ref{tab:point-likert-scores} list the average likert scores for the different points in each domain. In the map domain the points are named from top-left to bottom-right (e.g., $Pt1$ is the point in the top left of Figure \ref{fig:bottleneck-points}); in the classroom domain the points are named in order from left to right (e.g. $Pt1$ is the leftmost point marked in Figure \ref{fig:classromm-points}); and, in the military engagment domain the points are named from top to bottom (e.g., $Pt1$ is the point nearest the top of Figure \ref{fig:tanks-points}). For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli.

@@

\begin{table}
\centerline{
\begin{tabular}{l c c c l}
\hline
Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
Stim ID & P1 & P2 & P3 & Significance\\
(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & ToDo\\ % 1_1
(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & ToDo\\ % 1_2
\hline
Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
Stim ID & P1 & P2 & P3 & Significance\\
(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & ToDo\\ % 2_1
(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & ToDo\\ % 2_6
(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & ToDo\\ %2_7
(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & ToDo\\ %2_2
(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & ToDo\\ % 2_3
(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & ToDo\\ % 2_4
(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & ToDo\\\\ % 2_5
\hline
Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot\\
Stim ID & P1 & P2 & P3 & Significance\\
(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & ToDo\\ % 3_1
(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & ToDo\\ % 3_2
(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & ToDo\\ % 3_3
(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & ToDo\\ % 3_4
\hline
\hline
\end{tabular}
}
\label{tab:point-likert-scores}
\caption{Need to put in caption!!!}
\end{table}

-----------------

-----------------

Task 2: The participant selected a point denoting the sweetspot of the region
Analysis of results:
(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
We would hope that the variance of the subjects responses is lower than that of randomly generated points.
After the initial comparison with the random point sets we may exclude participants based on extreme variance.

(2) Does context matter?
Do the clusters of participants responses move?
- for each experimental stimulus compute the mean x and y location for the responses for that trial
- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.

-----------------

Task 3: The participant grades a point with respect to membership of the region on a likert scale
I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
(b) distance from the mean sweet spot for that stimulus from task 2

-----------------

>>>>

All three trial types use the same visual stimuli and accompanying spatial descriptions. Figure \ref{fig:exp_stimuli} presents the stimuli used in the experiment. Throughout the rest of the paper we will use the labels from Figure \ref{fig:exp_stimuli} as the names for the different stimuli. The stimuli can be broken into three groups based on the domain and the spatial region that accompanied them: the classroom domain stimuli were always accompanied by the spatial description \emph{the front of the classroom}; the geographical map stimuli were always accompanied by the description \emph{the bottleneck}; and the military engagement stimuli were always accompanied by the description \emph{safety}.

@@

\subsection{Participants and Procedure}

43 undergraduate students from two different third level institutes took part in the study. To control for sequence effects the presentation order of the trials were randomised each participant. Also in order to minimise fatigue of the subjects for the Type 2 Likert trials each subject only ranked a randomly selected subset of the trials.

@@

The type 2 trials the same visual stimuli and accompanying spatial descriptions as the type 1 trials with the addition that a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region. For each stimuli domain (classroom, map, military engagement) three points (P1, P2, and P3) were pre-selected.

The location of these points was informed by the results of a pilot study and the final design of the study and they were positioned such that for each particular stimuli one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region. In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}). One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped. In the classroom domain the expected sweetspot varied between the points depending on the stimuli. In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets. P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.

Table \ref{tab:point-likert-scores} lists the average likert scores for the different points in each domain. For each stimuli we have highlighted in bold the maximum average likert rating for a point for that stimuli. In all cases the expected sweetspot received the maximum average likert rating for trials using that stimuli. For each stimuli we tested whether the higher ratings of the sweetspot locations was statistically significant using one-tailed two sampl t-tests. There were two of these t-tests carried out for each stimuli, each one tested whether the sweetspot ranking was statistically higher than one of the other points. Table \ref{tab:point-likert-scores} also lists the results of these t-tests for each stimuli. In all cases, the results indicate that the null hypothesis, $H_0$, should be rejected and that the alternative hypothesis, namely that the sweetspot was rated significantly higher than the other point, could be accepted with $\alpha=0.05$.

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. Table \ref{tab:flipped-point-likert-scores} lists the results of this trial. In all cases the results indicated that there was not enough evidence to reject the null hypothesis, in other words there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli.

@@

\begin{table}
\centerline{
\begin{tabular}{l c c c l}
\hline
Domain & \multicolumn{4}{c}{Map/Bottleneck} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(g) & $1.1707$ & $\mathbf{4.6667}$ & $1.85$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_1
(h) & $1.1707$ & $\mathbf{4.7561}$ & $1.585$ & $H_A:\mu(P2)>\mu(P1),~p<2.2^{-16};~H_A:\mu(P2)>\mu(P3),~p<2.2^{-16}$\\ % 1_2
\\
\hline
Domain & \multicolumn{4}{c}{Classrom/Front of Room} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} & Sweetspot Rated Significantly Higher? \\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(a) & $\mathbf{3.5122}$ & $2.6341$ & $1.9024$ & $H_A:\mu(P1)>\mu(P2),~p=6.497^{-4};~H_A:\mu(P1)>\mu(P3),~p=4.073^{-8}$\\ % 2_1
(c) & $\mathbf{4.7073}$ & $2.9024$ & $1.3571$ & $H_A:\mu(P1)>\mu(P2),~p=6.396^{-16};~H_A:\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_6
(c-flipped) & $1.244$ & $2.9268$ & $\mathbf{4.6429}$ & $H_A:\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p:1.864^{-12}$\\ %2_7
(e) & $\mathbf{4.7561}$ & $2.0000$ & $1.2439$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3):~p<2.2^{-16}$\\ %2_2
(e-flipped) & $1.2439$ & $1.8500$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2):~p<2.2^{-16}$\\ % 2_3
(f) & $\mathbf{4.8292}$ & $2.0976$ & $1.2750$ & $H_A=\mu(P1)>\mu(P2),~p<2.2^{-16};~H_A=\mu(P1)>\mu(P3),~p<2.2^{-16}$\\ % 2_4
(f-flipped) & $1.1463$ & $2.3171$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<2.2^{-16}$\\ % 2_5
\\
\hline
Domain & \multicolumn{4}{c}{Military Engagement/Safe Region} \\
 ~ & \multicolumn{3}{c}{Average Likert Rating} &  Sweetspot Rated Significantly Higher?\\
Stim ID & P1 & P2 & P3 & One-Tailed Two Sample T-Test with $\alpha=0.05$, $H_0:\mu(P_i)=\mu(P_j)$\\
(i) & $1.1951$ & $4.3902$ & $\mathbf{4.9024}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=5.382^{-4}$\\ % 3_1
(j) & $1.0952$ & $2.2927$ & $\mathbf{4.5122}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p<7.981^{-14}$\\ % 3_2
(k) & $1.0714$ & $2.7804$ & $\mathbf{4.6585}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=2.708^{-13}$\\ % 3_3
(l) & $1.2327$ & $3.1667$ & $\mathbf{4.7073}$ & $H_A=\mu(P3)>\mu(P1),~p<2.2^{-16};~H_A=\mu(P3)>\mu(P2),~p=6.621^{-12}$\\ % 3_4
\\
\hline
\hline
\end{tabular}
}
\label{tab:point-likert-scores}
\caption{Average Likert Ratings per Point by Domain with T-Test results: in all cases the average rating of the sweetspot is significantly higher than the other points tested for that stimuli}
\end{table}

\begin{table}
\centerline{
\begin{tabular}{c c c c }
\hline
Domain & \multicolumn{3}{c}{Classrom/Front of Room} \\
$H_0$ & $H_A$ & $\alpha$ & p-value\\
$\mu(c.P1)=\mu(c-flipped.P3)$ & $\mu(c.P1)!=\mu(c-flipped.P3)$ & $0.05$ & $0.6315$\\
$\mu(e.P1)=\mu(e-flipped.P3)$ & $\mu(e.P1)!=\mu(e-flipped.P3)$ & $0.05$ & $0.7358$\\
$\mu(f.P1)=\mu(f-flipped.P3)$ & $\mu(f.P1)!=\mu(f-flipped.P3)$ & $0.05$ & $0.2592$\\
\hline
\hline
\end{tabular}
}
\label{tab:flipped-point-likert-scores}
\caption{Two Tailed Two Sample T-Tests for flipped stimuli pair trials.}
\end{table}

Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular situation depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted situation was consistent across similar situations, as shown by the fact that the sweetspot flipped from P1 to P3 when the situation was flipped.

%-----------------
%
%Task 2: The participant selected a point denoting the sweetspot of the region
%Analysis of results:
%(1) Do participants agree/understand-the-task? (validation of approach) + Should anyone be excluded?
%Compare the variance across the subjects with the variance across a number of set of randomly generate points, where set size equals the number of participants, using the F distribution (http://www.ltcconline.net/greenl/courses/201/regression/comparingVariances.htm) where we are checking if the variance of the random points are greater than that of the participants.
%We would hope that the variance of the subjects responses is lower than that of randomly generated points.
%After the initial comparison with the random point sets we may exclude participants based on extreme variance.
% 
%(2) Does context matter?
%Do the clusters of participants responses move?
%- for each experimental stimulus compute the mean x and y location for the responses for that trial
%- check whether the variance between the means of different stimuli responses is greater than the variance within each stimulus responses.

%-----------------
%
%
%Task 3: The participant grades a point with respect to membership of the region on a likert scale
%I understand this experiment to be useful because (a) it asks the participants to evaluate (rather than generate cf task 1 and 2) (b) hopefully we can use it to reinforce our findings in task 1 and task 2
%I think we need to be careful that the points we present to people for evaluation are reasonably well distributed so that we have a good chance of getting a variation across the responses
%Assuming that there is a variation across the likert scores for a particular stimulus then check if their is a correlation between the scores and:
%(a) inclusion in the intersection/union polygons from task 1 for the same stimulus
%(b) distance from the mean sweet spot for that stimulus from task 2
%
%-----------------
